Glue-on AdS holography for $T\bar T$-deformed CFTs

### 2023-08-24 alpha/GlueonAdS/glueon-poster-alt1.tex
Revamp glueon poster

\documentclass[11pt]{article}

@@

@@

\preauthor{\large}
\author{Wen-Xin Lai \textkai{赖文昕}}
\postauthor{, \textit{\footnotesize with}
	Luis Apolo,
	Peng-Xiang Hao \textkai{郝鹏翔}
	and Wei Song \textkai{宋伟}
	\ \arxiv{2303.04836}
	\par\vspace{.5\baselineskip}
	\hrule\vspace{.3\baselineskip}
}

@@

\section*{Foundation: pure Einstein in $\mrm{AdS}_3$~/~$\mrm{CFT}_2$}

\textcite{Maldacena:1997re}\\
\textcite{Aharony:1999ti}
\begin{align*}
	\textrm{Strings on $\mrm{AdS}_{d+1}$ background}
	&\ \equiv\ 
	\textrm{Conformal Field Theory $\mrm{CFT}_{d}$}
\\
	\textrm{\textit{asympt.}~$\mrm{AdS}_{d+1}$ Gravity}
	&\ \equiv\ 
	\textrm{Large $N$ $\mrm{CFT}_{d}$ at \textit{asympt.}~boundary}
\end{align*}
\begin{center}
	\includegraphics[width=.6\linewidth]{img/ads-cft.png}

	\vspace{-.2\baselineskip}
	\captionsetup{type=figure}
	\caption{$\mrm{AdS}_3/\mrm{CFT}_2$ cartoon}

	\vspace{.1\baselineskip}
	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{center}

@@

\item \textsc{Universal / ubiquitous}

\begin{itemize}
	\item Model agnostic: $\mrm{AdS}_5\times S^5$, $\mrm{AdS}_3\times S^3 \times T^4$, symmetric orbifolds, minimal models, ``monsters'', ... \textit{\small Maldacena, Witten, GKP, ABJM, Gaberdiel, and many many more}

@@

\item Our work: pure Einstein gravity in $\mrm{AdS}_3$ / $\mrm{CFT}_2$ $(d=2)$.\\
$\mrm{AdS}_3$ metric, $\rho \sim z^2 \sim r^{-2}$
\begin{equation}
\hspace{-2em}
\begin{aligned}
	ds^2
	&= \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j
\\[.5ex]
	&= \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)
	\label{fggauge}
\end{aligned}
\end{equation}

\vspace{-.5\baselineskip}
boundary at $\rho \to 0$, $x^{i} \sim \varphi, t$, indices $i,j = 1,2$

\item Surprisingly the expansion simply terminates at~$\rho\thinmspace g^{(4)}_{ij}$ for pure Einstein gravity! Moreover,
\begin{equation}
	g^{(4)}_{ij} = \frac{1}{4} g^{(2)}_{ik} g^{(0)\thinmspace kl}_{\phantom{kl}} g^{(2)}_{lj}
\end{equation}
\textcite{Fefferman:2007rka}\\
\textcite{Banados:1992wn}\\
\textcite{Banados:1998gg}
\end{itemize}

\newcommand{\eqBanados}{
	\begin{equation*}
	\hspace{-2em}
		ds^2 \ \,
	\end{equation*}
}

\begin{itemize}
\item $g^{(2)} \leadsto \mathcal L(u)$, $\bar{\mathcal L}(v)$: arbitrary periodic functions,\\
where $u,v = \varphi \mp it$, $du\,dv = d\varphi^2 + dt^2$
%	\vspace{-.5\baselineskip}

\item A consequence of Einstein equations \\
in $(2+1)$ dimensions.\\
\textit{Too many constraints, too little freedom!}
\end{itemize}

\begin{itemize}
\item Different \textbf{geometries} in the ``bulk''

specified by $\mathcal L(u)$, $\bar{\mathcal L}(v)$: periodic functions, maps to different \textbf{states} in the boundary $\mrm{CFT}_2$ with stress energy $\ave{T(u)}_\psi$, $\ave{\bar{T}(v)}_\psi$

\item \textit{Dictionary:} bulk $\leftrightarrow$ boundary quantities:
	\begin{itemize}
	\item spectrum,
		symmetry,\\
		partition functions,
		correlators, ...
	\end{itemize}
\end{itemize}

\begin{frame}{\TTbar deformations --- motivations}{%
	Comments taken from \textcite{Cui:2023jrb}
}
	\begin{itemize}
	\item Ideally: AdS/CFT
	\item Reality: non-AdS space and non-CFT
	\begin{itemize}
		\item possible to extend holographic principle to more general context?
		\item attempt: \textbf{\textit{deformations}} on both sides\\
		in a controllable way
	\end{itemize}
	\end{itemize}
\end{frame}

\newcommand{\citeTTbar}{%
	\textcite{Zamolodchikov:2004ce}, revived by \textcite{Smirnov:2016lqw} and \textcite{Cavaglia:2016oda} \textit{et al}
}

\begin{frame}{\TTbar deformations --- definition}{%
	\citeTTbar
}
\begin{itemize}

\item Define ``\,\TTbar\,'' as the flow of action:
\begin{equation}
	\partial_{\sidenote{\mu}} I = -{8 \pi} \int d^2x \, T\bar T =  -{\pi} \int d^2x\,\big( T^{ij}T_{ij}- (T^i_i)^2 \big)
	\label{TTbardef}
\end{equation}
For $\mrm{CFT}_2$,\ $T\bar{T} = T_{xx} T_{\bar{x}\bar{x}}$,\ where $x, \bar{x} = \varphi' \mp it'$.

\item The stress tensor $T_{ij}(\sidenote{\mu})$ on the right hand side flows with the deformation parametrized by $\sidenote{\mu}$.

\item This is thus a \textit{differential equation} for a flow, starting from some generic QFT specified by $I(\mu = 0)$. In this work we shall start from $\mrm{CFT}_2$, but note that \TTbar is \textit{irrelevant}: conformal symmetry will be broken.

\end{itemize}
\end{frame}

\begin{frame}{``Solvable'' deformations on the field theory side}{%
	\citeTTbar
}
\begin{itemize}

\item \textbf{``Solvable'':} the deformed spectrum of $\hat{H}(\mu)$ and $\hat{J}(\mu)$ on a cylinder of radius $R$ can be solved, 
and it's a simple function\\
 of the undeformed $E(0)$, $J(0)$:
\begin{equation}
\hspace{-2em}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0) % \label{ttbarspectrum}
\end{equation}
under simple (and reasonable) assumptions (e.g.~translation invariance et al).

\item \textbf{Variants:} suppose the theory has some ``components'' labeled by $w$,

\begin{itemize}
\item double-trace $\TTbar = (\sum_w T_w) (\sum_w \bar{T}_w)$ (this work)
\item single-trace $\TTbar' = \sum_w (T_w \bar{T}_w)$ (e.g.~\textit{\citeauthor{Cui:2023jrb}},~23)
\end{itemize}

@@

\vspace{.5\baselineskip}
\end{frame}

\begin{frame}{\TTbar deformations --- properties}{%
	\textcite{Zamolodchikov:2004ce}\\
	\citetitle{Zamolodchikov:2004ce}
}
\vspace{-.1\baselineskip}
\begin{equation}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \ \quad J(\mu)=J(0)  \label{ttbarspectrum}
\end{equation}
\vspace{-\baselineskip}
\begin{itemize}
\item It seems that the \TTbar deformation is well-defined non-perturbatively for generic $\mrm{QFT}_2$,
yet it's still mysterious and difficult\\
\begin{itemize}
\item e.g.~correlators: \textcite{Kraus:2018xrn,Cardy:2019qao,Cui:2023jrb}
\item ``asymptotic fragility'': \textcite{Dubovsky:2017cnj}
\end{itemize}

\item What if we start from a $\mrm{CFT}_2$, that enjoys a holographic dual to $\mrm{AdS}_3$...

\end{itemize}
\end{frame}

\newcommand{\figCutoffAds}[1][.5]{
\begin{center}
	\hspace{.15\linewidth}
	\includegraphics[width=#1\linewidth]{img/cutoff-ads.jpg}

	\vspace{-.5\baselineskip}
	\ \scriptsize\textcite{AldegundePWSep22}
\end{center}
}

\begin{frame}{Cutoff $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}

\figCutoffAds

\begin{itemize}
\item Question: what is the AdS dual of \TTbar deformation?

\begin{itemize}
\item look up \textit{double-trace deformations}\\
in \textit{the dictionary}:\\
\textcite{Heemskerk:2010hk}

\item look at properties of \TTbar:\\
signs of gravity and coarse-graining\\
\textcite{Dubovsky:2012wk,Dubovsky:2013ira}
\end{itemize}

\item Answer: AdS gravity within a finite Dirichlet wall

@@

\end{frame}

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{}

	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}
	\textit{Roughly $\mu \sim 1/r^2$}

	\item Passed many non-trivial tests,\\
	related to ``mixed boundary conditions''\\
	by \textcite{Guica:2019nzm}

\end{frame}

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}

\begin{itemize}
\item \textbf{Significance:} gravity inside a finite box\\
	with hard Dirichlet walls

\begin{itemize}
	\item Dual to some ``solvable'' no-longer-conformal ft
	\item A step towards quantum gravity in reality!
\end{itemize}

\item \textbf{Caveat:} the duality only admits $\zeta_c > 0$ so $\mu < 0$
\begin{equation}
	\zeta_c = - \frac{c \mu}{3\ell^2}
%		\label{dictionary}
\end{equation}
But \TTbar itself admits $\mu > 0$ with nice properties.\\
\textit{What is the other side of the duality?}
	\begin{itemize}
	\item For comparison, the proposal of \textcite{Guica:2019nzm} admits both signs of $\mu$
	\end{itemize}
\end{itemize}

\end{frame}

\section{\textbf{Proposal:} Glue-on AdS holography} \label{se:glueonproposal}

\begin{frame}{Glue-on $\mrm{AdS}_3$ --- analytic continuation}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric has only simple poles with the $\rho$ coordinate:
	\begin{equation}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ %\label{banados}
	\end{equation}
	thus it admits well-defined analytic continuation \\
	from $\rho > 0$ to $\rho \in \mbb{R}$.
	\item Continuation of the dictionary:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
	\ \quad
		\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
	\end{equation}
%	\vspace{-.5\baselineskip}

	\end{itemize}
\end{frame}

\newcommand{\figGlueon}[1][1]{
\begin{minipage}{#1\linewidth}
	\centering
	\includegraphics[width=.5\linewidth]{img/diagram.pdf}
	\vspace{-1.5\baselineskip}
	\captionsetup{type=figure}
	\caption{Glue-on $\mrm{AdS}_3$}

	\vspace{-.3\baselineskip}
	\scriptsize\ Top-down view of the Poincar\'e disk
\end{minipage}
}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}

\figGlueon

\begin{equation}
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
\ \quad
	\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
\end{equation}
\begin{itemize}
\item What have we done other than copy-pasting?
	\begin{itemize}
	\item Although the continuation is straightforward, physical quantities diverge as $\rho \to 0$

	\item A prescription is required to ``renormalize'' the divergences
	\end{itemize}

\end{frame}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Kraus:2018xrn} -- \citetitle{Kraus:2018xrn}
}
\begin{itemize}
\item Firstly, matching energy momentum (Brown-York) \& the flow equations:
	\begin{equation}
		T_{ij}= \frac{\sigma_{\zeta_c}}{8\pi G} \Big( K_{ij} -  K \gamma_{ij} + \frac{1}{\ell |\zeta_c|}\gamma_{ij}\Big), \ \quad \sigma_{\zeta_c} = \frac{\abs{\zeta_c}}{\zeta_c} = \pm 1,
		\label{BYtensor}
	\end{equation}
	The field theory metric $\gamma_{ij} = \zeta_c h_{ij}$, where $h_{ij}$ is the induced metric.

	\begin{itemize}
	\item $\gamma_{ij}$ is always positive-definite, while $h_{ij}$ becomes negative-definite for the glue-on region.
	This discrepancy is the origin of the $\abs{\zeta_c}$.
	\end{itemize}

\item \TTbar flow is recast geometrically as the $i,j$ components of the Einstein equations. 
Note that the geometry can be decomposed such that:
	\begin{equation}
		ds^2 = \frac{1}{\zeta} \gamma_{ij}dx^i dx^j+n_\mu n_\nu dx^\mu dx^\nu
	\end{equation}
\end{itemize}
\end{frame}

\begin{frame}{Glue-on $\mrm{AdS}_3$ --- beyond the infinity}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{itemize}
\item The geometry is foliated by constant $\zeta$ surfaces ${\mathcal N}_\zeta$:
	\begin{equation}
	\begin{aligned}
		ds^2 
		&= \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg) \\
		&= \frac{1}{\zeta} \gamma_{ij}dx^i dx^j+n_\mu n_\nu dx^\mu dx^\nu,\ \quad
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi}
	\end{aligned}
	\end{equation}
	The \TTbar deformed theory lives on $\mcal{N}_{\zeta_c}$ with metric $\gamma_{ij}$.

\item Treatment for the singularity at $\zeta\to 0^\pm$: introduce $\mathcal N_{\zeta={\pm\epsilon}}$ and glue them together (``glue-on'');
	exclude the $-\epsilon < \zeta < \epsilon$ region until finally sending $\epsilon \to 0$. 
\item We formally denote the boundary surface by the limit $\mathcal N_{0}=\mathcal N_{0^+}=\mathcal N_{0^-}$, though we need to keep track of the asymptotic cutoff $\epsilon$ in actual computations.
\end{itemize}
\end{frame}

\newcommand{\stateGlueon}{\ensuremath{
	\textrm{Cutoff / \textit{glue-on} $\mrm{AdS}_{d+1}$ Gravity}
	\ \equiv\ %
	\textrm{\TTbar deformed $\mrm{CFT}_{d}$ at $\mcal{N}_{\zeta_c}$}
}}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- the proposal}{}

\begin{center}
\includegraphics[width=.6\linewidth]{img/diagram.pdf}
\begin{equation*}
\stateGlueon
\end{equation*}
\end{center}

\end{frame}

\section{\textbf{Check:} \TTbar deformed charges \& partition functions} \label{se:charges}

\begin{frame}{\TTbar quantities from bulk geometry}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}

\begin{itemize}
\item AdS/CFT: weak / strong duality
	\begin{itemize}
	\item \textit{quantum} quantities on the CFT side\\
		can be computed by (semi-)classical geometry
	\item inherited by the \TTbar deformation
	\end{itemize}

\item Demanding the extended geometry to be non-singular
	re-produces bounds on the \TTbar deformed theory, e.g.
	\begin{equation}
		\zeta_c \ge -1 \quad \Leftrightarrow \quad\mu\le \frac{ 3\ell^2 }{c} \label{reality}
	\end{equation}
	$\zeta_c = -1$ is a horizon of the glue-on geometry, \\
	where $\det g_{\mu\nu} = 0$.

\end{frame}

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}

\begin{itemize}
\item Spectrum: conserved charges of geometries,\\
	e.g.~black holes \textit{BTZ} \cite{Banados:1992wn}
	\begin{itemize}
	\item computed with the covariant formalism\\
		\textcite{Iyer:1994ys, Barnich:2001jy}
	\item manifest covariance (diff-invariance) depends on reparametrization redundancies (gauge),\\
		so delay gauge fixing until the very end
	\end{itemize}

\item We are careful to match the bulk \& boundary symmetries at $\mcal{N}_{\zeta_c}$

\item This reproduces $E(\mu)$, $J(\mu)$ with $\mu \in\mbb{R}$ in \eqref{ttbarspectrum}

\end{frame}

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}

\begin{itemize}
\item Covariant charges of geometries
	\begin{align*}
		\delta E(\mu) &\equiv  \ell^{-1} \delta \mathcal Q_{\pd_{t'}}, \\ \delta J(\mu) &\equiv  \delta \mathcal Q_{\pd_{\varphi'}},
	\end{align*}

\item One needs to choose the appropriate boundary coordinates $(t',\varphi')$, such that
	\begin{equation}
		ds^2_c = \ell^2 ( -dt'^2 + d\varphi'^2) , \qquad \varphi' \sim \varphi' + 2\pi.\label{cutoffmetric}
	\end{equation}

\item This is realized by the \textit{state-dependent} map of coordinates
	\begin{align*}
		dt' &= \sqrt{ \big(1 - \zeta_c (T_u+T_v)^2 \big) \big(1 - \zeta_c (T_u-T_v)^2 \big) } \, dt,  \\
		 d\varphi' &= d\varphi + \zeta_c \big( T_u^2 - T_v^2 \big)\,dt.
	\end{align*}
\end{itemize}

\end{frame}

\begin{frame}{\TTbar thermodynamics from bulk geometry}{%
	\textcite{Giveon:2017nie,Apolo:2019zai}
}

@@

\item Many consequences: correct charges, signal propagation speed $v'_{\pm} \equiv \pm \frac{d\varphi'}{dt'}$, and thermodynamics when Wick rotated:
	\begin{equation}
		T_L(\mu)\,T_R(\mu) \le - \frac{1}{4\pi^2 \ell^2 \zeta_c} = \frac{3}{4\pi^2c\mu} = T_H(\mu)^2,
	\quad \mu > 0
	\end{equation}
	$T_{L,R}$ are temperatures associated with $u',v' = \varphi' \pm t'$.

\item $T_H$ is the \textit{Hagedorn temperature}: exceeding $T_H$ corresponds to a complex entropy
\end{itemize}

\end{frame}

\section{\textbf{Check:} \TTbar partition functions from glue-on AdS} \label{se:partitionfunction}

\begin{frame}{\TTbar partition functions from bulk on-shell actions}{%
	\textcite{Caputa:2020lpa}
}

\begin{itemize}
\item Bulk: weakly coupled gravity, the partition function is approximated by the on-shell gravity action of the dominant saddle:
	\begin{equation}
		Z_{T\bar T} (\mu) = \mathcal Z (\zeta_c) \approx  e^{-I(\zeta_c)}\label{partition2}
	\end{equation}

\item $I(\zeta_c)$ is roughly the volume of the geometry, which diverges at $\zeta \to 0^\pm$. We apply a natural extension of \textit{holographic renormalization} to remove the divergence.

\item $I = -\log \mcal{Z}$ satisfies the \TTbar flow \eqref{TTbardef}.\\
	This is non-trivial because we've packaged the quantum corrections of the boundary theory.
\end{itemize}
\end{frame}

\begin{frame}{Comparison with field theory analysis}{%
	\textcite{Datta:2018thy,Apolo:2023aho}
}

\vspace{-.3\baselineskip}

\begin{itemize}
\item \textbf{Torus:} Large $c$, \textit{modular invariance} \& sparseness of the ``light'' spectrum \cite{Datta:2018thy,Apolo:2023aho},
	c.f. \textcite{Hartman:2014oaa}
	\begin{equation*}
	\hspace{-1.5em}
		\log   Z_{T\bar T}(\mu)  \approx \left\{ \begin{aligned}
		& {-\frac{1}{2}}\,(\beta_L + \beta_R)\, RE_{\text{vac}}(\mu),  &\beta_L \beta_R > 1, \\
		& {-2 \pi^2 \bigg(\frac{1}{\beta_L} + \frac{1}{\beta_R}\bigg)  RE_{\text{vac}}\bigg(\frac{4\pi^2}{\beta_L \beta_R} \mu \bigg)}, \ \, &\beta_L\beta_R < 1, %\\[2ex]
		 \end{aligned} \right. %\label{ZTTbar}
	\end{equation*}

\item \textbf{Sphere:} maximally symmetric, \textcite{Donnelly:2018bef}
	\begin{equation}
%		\label{sphere:stresstensor}
		\langle T_{ij}\rangle =-\frac{1}{4\pi\mu} \bigg(1-\sqrt{1-\frac{c\mu}{3L^2}}\, \bigg)  \gamma_{ij}.
	\end{equation}
\end{itemize}
\end{frame}

>>>>

\documentclass[10pt]{article}

@@

\newcommand{\citations}[1]{{\footnotesize#1\par}}
\newcommand{\reality}{reality\textsuperscript{TM}}

@@

\preauthor{\par\large\noindent}
\author{Wen-Xin Lai \textkai{赖文昕}}
\postauthor{, \textit{\normalsize with}
	Luis Apolo,
	Peng-Xiang Hao \textkai{郝鹏翔}
	and Wei Song \textkai{宋伟}
	\hfill\arxiv{2303.04836}
	\par\vspace{.5\baselineskip}
	\hrule\vspace{.3\baselineskip}
}

@@

\subsection*{Foundations:\texstringonly{\\}Pure Einstein gravity in $\mrm{AdS}_3$~/~$\mrm{CFT}_2$}

\noindent\citations{%
\textcite{Fefferman:2007rka}\\
\textcite{Banados:1992wn}\\
\textcite{Banados:1998gg}
}

We focus on pure Einstein gravity without matter in the bulk; the Einstein equations are solved by the metric:
\begin{equation}
\hspace{-2em}
\begin{aligned}
	ds^2
	&= \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\\
	&= n_\mu n_\nu dx^\mu dx^\nu + \frac{1}{\zeta} \gamma_{ij}dx^i dx^j,\quad
\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \equiv r^2
\end{aligned}
\label{fggauge}
\end{equation}
The radial coordinate: $\rho \approx r^{-2} \equiv \zeta$,
where $r$ is the ``proper radius''; $\mathcal L(u)$ and $\bar{\mathcal L}(v)$ are arbitrary periodic functions.

\noindent%
The geometry is foliated by constant $\zeta$ surfaces ${\mathcal N}_\zeta$. 
\\
The asymptotic boundary $\mcal{N}_0$ is at $\zeta \to 0$, with coordinates: $x^{i} \sim \varphi, t$, while $u,v = \varphi \pm t$.

\noindent%
Our work is based on the $\mrm{AdS}_3/\mrm{CFT}_2$ duality:
\begin{align*}
%\hspace{-1em}
	\textrm{\textit{asympt.}~$\mrm{AdS}_3$ Gravity}
	&\ \equiv\ 
	\textrm{Large $N$ $\mrm{CFT}_2$ at $\mcal{N}_0$}
\end{align*}
\citations{\noindent%
	\textcite{Maldacena:1997re}\\
	\textcite{Aharony:1999ti} \textit{et al}
}
%\begin{center}
%	\includegraphics[width=.7\linewidth]{img/ads-cft.png}\\[.5ex]
%	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
%\end{center}

%\begin{itemize}

%\item Universal / ubiquitous:
%
%\begin{itemize}
%	\item Model agnostic: $\mrm{AdS}_5\times S^5$, $\mrm{AdS}_3\times S^3 \times T^4$, symmetric orbifolds, minimal models, ``monsters'', ... \textit{\small Maldacena, Witten, GKP, ABJM, Gaberdiel, and many many more}
%	
%\end{itemize}

%\item

%\end{itemize}

\subsection*{\TTbar deformations \& cutoff $\mrm{AdS}_3$}

\citations{
\textcite{Zamolodchikov:2004ce}\\
\textcite{Dubovsky:2012wk}\\
\textcite{Dubovsky:2013ira}\\
\textcite{Smirnov:2016lqw}\\
\textcite{Cavaglia:2016oda}\\
\textcite{Dubovsky:2017cnj} \textit{et al}
}

@@

\item Define ``\,$\TTbar$\,'' as the flow of action:
\begin{equation}
	\partial_{\sidenote{\mu}} I = {8 \pi} \int d^2x \, T\bar T_{(\sidenote{\mu})} = {\pi} \int d^2x\,\big( T^{ij}T_{ij}- (T^i_i)^2 \big)_{(\sidenote{\mu})}
	\label{TTbardef}
\end{equation}
where $x, \bar{x} = \varphi' \pm t'$.
We shall start from $\mrm{CFT}_2$ such that $T\bar{T}_{(\mu = 0)} = T_{xx} T_{\bar{x}\bar{x}}$, but note that the deformation is \textit{irrelevant}: conformal symmetry will be broken for $\mu \ne 0$.

\item \textbf{Solvable:} the deformed spectrum of $\hat{H}(\mu)$ and $\hat{J}(\mu)$ on a cylinder of radius $R$ is a simple function:
\begin{equation}
\hspace{-2em}
\begin{gathered}
	E(\mu) = - \frac{R }{ 2\mu } \bigg(1-\sqrt{1 + \frac{4\mu}{R} E(0) + \frac{4\mu^2}{R^4} J(0)^2 }
	\,\bigg), \\[1ex] J(\mu)=J(0)
\end{gathered} \label{ttbarspectrum}
\end{equation}
of the undeformed spectrum $E(0)$, $J(0)$,
under reasonable assumptions (e.g.~translation invariance).

@@

\subsection*{Cutoff $\mrm{AdS}_3$ duality:\texstringonly{\\} Holography within a finite Dirichlet wall}

\citations{
\textcite{McGough:2016lol}\\
\textcite{Kraus:2018xrn} \textit{et al}
}\vspace{-.8\baselineskip}
\begin{center}
	\hspace{.1\linewidth}
	\includegraphics[width=.48\linewidth]{img/cutoff-ads.jpg}

	\vspace{-.5\baselineskip}
	\scriptsize\ Image courtesy: \textcite{AldegundePWSep22}
\end{center}
\textbf{Dictionary:} the \TTbar deformed theory lives on $\mcal{N}_{\zeta_c}$ with metric $\gamma_{ij}$.
Location (radius) of the cutoff surface gets mapped to the deformation parameter $\mu$:
\begin{equation}
	\zeta_c = - \frac{c \mu}{3\ell^2}
	\label{dictionary}
\end{equation}
\TTbar flow is recast geometrically as the $i,j$ components of the Einstein equations.

\columnbreak

\item \textbf{Significance:} gravity inside a finite box, dual to some ``solvable'' no-longer-conformal theory.

\textit{A step towards quantum gravity in \reality!}

\item \textbf{Caveat:} the duality only admits $\zeta_c > 0$ so $\mu < 0$.
But \TTbar itself admits $\mu > 0$ with nice properties.

For comparison, the related proposal of \textcite{Guica:2019nzm} admits both signs of $\mu$.

\textit{What is the other side of the duality?}

@@

\section*{\textbf{Proposal:}\texstringonly{\\}Glue-on $\mrm{AdS}_3$ -- beyond the infinity}
\label{se:glueonproposal}

\parbox{\linewidth}{
\vspace{-\baselineskip}
\begin{equation*}
	\textrm{Cutoff / \textit{glue-on} $\mrm{AdS}_{d+1}$ Gravity}
	\ \equiv\ %
	\textrm{\TTbar deformed $\mrm{CFT}_{d}$ at $\mcal{N}_{\zeta_c}$}
\end{equation*}
\begin{center}
	\vspace{-.8\baselineskip}%
	\centering
	\includegraphics[width=.6\linewidth]{img/diagram.pdf}

	\vspace{-.5\baselineskip}
	\scriptsize\ Top-down view of a constant $t$ slice
\end{center}
%\vspace{-\baselineskip}
}

\item \textbf{Analytic continuation:} metric \eqref{fggauge} has only simple poles, thus admits well-defined analytic continuation:
%\begin{equation}
%	\rho,\ \zeta,\ \mu\ \in\ \mbb{R}
%\end{equation}
\begin{equation}
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
\ \quad
	\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
\end{equation}
What are we doing other than copy-pasting?
	\begin{itemize}
	\item Physical quantities diverge as $\rho \to 0$.
	A prescription is required to ``renormalize'' the divergences.

	\item Treatment for the singularity at $\zeta\to 0^\pm$: introduce $\mathcal N_{\zeta={\pm\epsilon}}$ and glue them together (``\textbf{glue-on}'');
		exclude the $-\epsilon < \zeta < \epsilon$ region until finally sending $\epsilon \to 0$. 
	\item We formally denote the boundary surface by the limit $\mathcal N_{0}=\mathcal N_{0^+}=\mathcal N_{0^-}$, though we need to keep track of the asymptotic cutoff $\epsilon$ in actual computations.
	\end{itemize}

\item \textbf{Extending the flow:} matching energy-momentum (Brown-York) \& the \TTbar flow equations:
	\begin{equation}
	\begin{gathered}
		T_{ij}= \frac{\sigma_{\zeta_c}}{8\pi G} \Big( K_{ij} -  K \gamma_{ij} + \frac{1}{\ell |\zeta_c|}\gamma_{ij}\Big), \\ \sigma_{\zeta_c} = \frac{\abs{\zeta_c}}{\zeta_c} = \pm 1,\\
		\gamma_{ij} = \zeta_c h_{ij},
	\end{gathered}	\label{BYtensor}
	\end{equation}
	where $h_{ij}$ is the induced metric.

	The field theory metric $\gamma_{ij}$ is always positive-definite, while $h_{ij}$ becomes negative-definite for the glue-on region $\rho,\zeta < 0$.
	This discrepancy is the origin of the $\abs{\zeta_c}$.

@@

\subsection*{\TTbar physics from the glue-on geometry}

@@

\item Demanding the extended geometry to be non-singular
	re-produces bounds on the \TTbar deformed theory, e.g.
	\begin{equation}
		\zeta_c \ge -1 \quad \Leftrightarrow \quad\mu\le \frac{ 3\ell^2 }{c} \label{reality}
	\end{equation}
	This is a Killing horizon of the glue-on geometry;
	$(\det g_{\mu\nu})_{\,\zeta = -1} = 0$.

\item \textbf{Spectrum from the extended geometry:} \mbox{conserved} charges of the Killing symmetries, can be computed with the \textit{covariant formalism}:

\citations{
	\textcite{Iyer:1994ys}\\
	\textcite{Barnich:2001jy}\\
	With \TTbar: \textcite{Kraus:2021cwf}
}\vspace{-1.5\baselineskip}
	\begin{align*}
		\delta E(\mu) &\equiv  \ell^{-1} \delta \mathcal Q_{\pd_{t'}}, \\ \delta J(\mu) &\equiv  \delta \mathcal Q_{\pd_{\varphi'}},
	\end{align*}
	This reproduces $E(\mu)$, $J(\mu)$ with $\mu \in\mbb{R}$ in \eqref{ttbarspectrum}.

\item We are careful to match the bulk \& boundary symmetries at $\mcal{N}_{\zeta_c}$.
One needs to choose the appropriate boundary coordinates $(t',\varphi')$, such that
	\begin{equation}
		ds^2_c = \ell^2 ( -dt'^2 + d\varphi'^2) , \qquad \varphi' \sim \varphi' + 2\pi.\label{cutoffmetric}
	\end{equation}
This is realized by:
	\begin{align*}
		dt' &= \sqrt{ \big(1 - \zeta_c (T_u+T_v)^2 \big) \big(1 - \zeta_c (T_u-T_v)^2 \big) } \, dt,  \\
		 d\varphi' &= d\varphi + \zeta_c \big( T_u^2 - T_v^2 \big)\,dt.
	\end{align*}

\item The above \textbf{\textit{state-dependent} map of \mbox{coordinates}} leads to the correct charges, the modified signal propagation speed $v'_{\pm} \equiv \pm \frac{d\varphi'}{dt'}$, and the \TTbar \textbf{thermodynamics} when Wick rotated. In particular,
	\begin{equation}
	\hspace{-2.5em} \mu > 0,\ \quad
		T_L(\mu)\,T_R(\mu) \le - \frac{1}{4\pi^2 \ell^2 \zeta_c} = \frac{3}{4\pi^2c\mu} = T_H(\mu)^2.
	\end{equation}
	$T_{L,R}$ are temperatures associated with $u',v' = \varphi' \pm t'$.\\
	$T_H$ is the \textit{Hagedorn temperature}: exceeding $T_H$ corresponds to a complex entropy.

\citations{
\textcite{Giveon:2017nie}\\
\textcite{Apolo:2019zai}
}
\end{itemize}

\subsection*{\TTbar partition functions from bulk actions} \label{se:partitionfunction}

\citations{
\textcite{Caputa:2020lpa}
}
